### 2023-09-09 manuscript_supp.Rmd

# Classification of vegetation into vertical strata

Regardless of taxonomic and functional identity, plants of height $\le$ 1 m above the ground classified into the 'Herb' layer (this therefore included both woody- [like Rubus] and non-woody stemmed [like graminoids] plants, but did not exclude graminoid leaves that extended beyond 1 m from the ground); plants 1--4 m in height belonged to the 'Understorey' (this included shrubs like S. aucuparia and F. alnus as well as young individuals of trees like P. abies); and plants $\ge$ 4 m in height belonged to the 'Canopy' (this occasionally included S. aucuparia).

Counts of 'Canopy' and 'Understorey' plants were taken and used to calculate density per 100 m\^2. Counts were not taken for the 'Herb' layer due to the difficulty in some cases in determining what constitutes a single stem and since it is unlikely that birds track ground layer vegetation at the level of individual stems. Instead, the species that covered $\gt$ 50% of the ground was considered to be the 'Dominant Ground Vegetation' (in some cases, this was moss) and this variable was also used in analyses.

# Phenological measurements

-   Table showing the ordinal scale of each variable

-   OR, possible options in each variable

# Table 2: all measured variables

Both habitat and bird count variables

continuous/ordinal/categorical, correlation, exclusion, random/fixed effect

# Vegetation community composition

As such, the tree community consists mostly of conifers like Norway spruce, Scots pine *Pinus sylvestris*, European larch *Larix decidua* and silver fir *Abies alba*, and broad-leaved species such as pedunculate/European oak *Quercus robur*, sessile/Cornish oak *Q. petraea*, northern red oak *Q. rubra*, silver birch *Betula pendula*, quaking/Eurasian aspen *Populus tremula*, common/European beech *Fagus sylvatica*, rowan or European mountain ash *Sorbus aucuparia*, small-leaved lime or littleleaf linden *Tilia cordata*, common/European black alder *Alnus glutinosa*, Norway maple *Acer platanoides* and sycamore maple *A. pseudoplatanus*. The commonly found shrub species are alder buckthorn *Frangula alnus*, rowan *Sorbus aucuparia* and black/European elderberry *Sambucus nigra*, while ground layer vegetation is restricted to graminoids like sedges *Carex* spp. (mostly *C. brizoides*), bramble *Rubus* sp. and European blueberry *Vaccinium myrtillus*.