### 2023-09-16 manuscript/03_title.Rmd
updated rmd structure to bookdown2

---
title: "..."
author: 
  - Karthik Thrikkadeeri:
      email: rikudoukarthik@gmail.com
      institute: usb
      correspondence: true
  - name: 
    institute: 
institute:
  - usb: 'University of South Bohemia, Branišovská 1645/31A, České Budějovice 2, 370 05 České Budějovice, Czechia'
output: 
  bookdown::word_document2:
    number_sections: false
    reference_docx: ../rmd_word_template_journ_guide.docx
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
---

# Acknowledgements

TBD

# Conflict of interest

The authors declare that they have no conflicting financial or personal interests that could have influenced this study.

# Supplementary Online Material (SOM)

More details about the methodology, analyses and results are available online in supplementary materials.